### 2022-07-07 bools/bools.tex
martin suda comments 2/2

\looseness=-1
A {\em signature} is a quadruple $(\Sigmaty, \Vty, \Sigma, \VV)$ where $\Sigmaty$ is a
set of type constructors, $\Vty$ is a set of type variables and $\Sigma$ and
$\VV$ are sets of constants and term variables, respectively. We require nullary
type constructor $o$ as well as binary constructor $\rightarrow$
to be in $\Sigmaty$. A type $\tau, \upsilon$ is either a type variable $\alpha \in
\Vty$ or of the form $\kappa(\tau_1, \ldots \tau_n)$ where $\kappa$ is an
$n$-ary type constructor. We write $\kappa$ for $\kappa()$, $\tau \rightarrow
\upsilon$ for $\rightarrow(\tau, \upsilon)$, and we drop parentheses to shorten 
$\tau_1 \rightarrow (\cdots \rightarrow (\tau_{n-1} \rightarrow \tau_n) \cdots)$ into $\tau_1 \rightarrow \cdots \rightarrow
\tau_n$. Each symbol in $\Sigma$ is
assigned a type declaration of the form $\forallty{\tuplen{\alpha}} \tau$ where all variables
occurring in $\tau$ are among $\tuplen{\alpha}$.

@@

\newcommand{\eqneq}{\mathrel{\dot{\eq}}}
These two rules hoist Boolean subterms $f$ to the literal level; therefore,
some results of \infname{Cases} and \infname{CasesSimp} will have literals of the form $f \eq \itrue$ (or
$f \noteq \itrue$) where $f$ is not an atom. This introduces the need for the rule
called immediate clausification (\infname{IC}), where $\tuple{D}{m}$ are defined below:
%
$$ \namedsimp{IC}{C}{D_1 \; \cdots \; D_m} $$
%
Let $s \eqneq t$ denote $\eq$ or $\not\eq$.
We say that a clause is \emph{standard} if all of its literals are of the form $s \eqneq t$,
where $s$ and $t$ are not Booleans or of the form $f \eqneq \itrue$, where the head of $f$
is not a logical symbol. The rule \infname{IC}
is applicable if clause $C = L_1 \llor
\cdots \llor L_n$ is not standard.
The resulting clauses $\tuple{D}{m}$ represent
the result of clausification of the formula $\iforall\,\tuple{x}{}. \; L_1 \ior
\cdots \ior L_n$ where $\tuple{x}{}$ consists of all free variables of $C$.
Using Boolean extensionality, Zipperposition's clausification
algorithm treats Boolean equality as equivalence (i.e., it replaces
$\mathop{\eq}\typeargs{o}$ with $\iequiv$).

@@

Expressiveness of higher-order logic allows users to define equality using a single axiom,
called Leibniz equality \cite{pa-01-classical-ty-thy}:
%
$ \iforall xy. \,
   (\iforall p. \, p \, x \iimplies p \, y) \iimplies x \ieq y$.
%
Intuitively, it embodies the principle that if two terms are indistinguishable by propositions,
they must be equal.
Leibniz equality often appears in TPTP problems  \cite{gs-17-tptp}. Since modern provers have native support
for equality, it is usually beneficial to recognize and replace occurrences of Leibniz equality.

@@

\looseness=-1
Combining these two rules with dynamic clausification is more useful: When rule
\infname{IC} is replaced by rule \infname{DC}, the success rate increases
to 1660 problems, compared to 1646 problems solved by b. We also discovered
that reasoning with choice is useful: When rule \infname{Choice} is enabled, the
success rate increases to 1653. We determined that including or excluding the
conclusion $D$ of \infname{Choice}, after it is simplified, makes no difference.
Counterintuitively, disabling the \infname{BoolSimp} rule results in 1640 problems,
which is only 6 problems short of configuration b. Disabling the \infname{Abs} and
\infname{Interpret$\lambda$} rules results in solving 25 and 31 problems less,
respectively. The raw data show that in total, using configurations from Figure
\ref{fig:cases} to Figure \ref{fig:bool-rules}, 1682 problems can be solved.

Last, we compared our approach to alternatives. Axiomatizing Booleans brings
Zipperposition down to a grinding halt: only 1106 problems can be solved using
this mode. On the other hand, preprocessing is fairly competitive: it solved
only 8 problems less than the b configuration.

>>>>

\looseness=-1
A {\em signature} is a quadruple $(\Sigmaty, \Vty, \Sigma, \VV)$ where $\Sigmaty$ is a
set of type constructors, $\Vty$ is a set of type variables and $\Sigma$ and
$\VV$ are sets of constants and term variables, respectively. We require nullary
type constructor $o$ as well as binary constructor $\rightarrow$
to be in $\Sigmaty$. Types $\tau, \upsilon$ are either type variables $\alpha \in
\Vty$ or of the form $\kappa(\tau_1, \ldots \tau_n)$ where $\kappa$ is an
$n$-ary type constructor. We write $\kappa$ for $\kappa()$, $\tau \rightarrow
\upsilon$ for $\rightarrow(\tau, \upsilon)$, and we drop parentheses to shorten 
$\tau_1 \rightarrow (\cdots \rightarrow (\tau_{n-1} \rightarrow \tau_n) \cdots)$ into $\tau_1 \rightarrow \cdots \rightarrow
\tau_n$. Each symbol in $\Sigma$ is
assigned a type declaration of the form $\forallty{\tuplen{\alpha}} \tau$ where all variables
occurring in $\tau$ are among $\tuplen{\alpha}$.

@@

\newcommand{\eqneq}{\mathrel{\dot{\eq}}}
These two rules hoist Boolean subterms $f$ to the literal level; therefore,
some conclusions of \infname{Cases} and \infname{CasesSimp} will have literals of the form $f \eq \itrue$ (or
$f \noteq \itrue$) where $f$ is not an atom. This introduces the need for the rule
called immediate clausification (\infname{IC}), where $\tuple{D}{m}$ are defined below:
%
$$ \namedsimp{IC}{C}{D_1 \; \cdots \; D_m} $$
%
Let $s \eqneq t$ denote $\eq$ or $\not\eq$.
We say that a clause is \emph{standard} if all of its literals are of the form $s \eqneq t$,
where $s$ and $t$ are not Booleans or of the form $f \eqneq \itrue$, where the head of $f$
is not a logical symbol. The rule \infname{IC}
is applicable if clause $C = L_1 \llor
\cdots \llor L_n$ is not standard.
The resulting clauses $\tuple{D}{m}$ represent
the result of clausification of the formula $\iforall\,\tuple{x}{}. \; L_1 \ior
\cdots \ior L_n$ where $\tuple{x}{}$ consists of all free variables of $C$.
Using Boolean extensionality, Zipperposition's clausification
algorithm treats Boolean equality as equivalence (i.e., it replaces
$\mathop{\eq}\typeargs{o}$ with $\iequiv$).

@@

Expressiveness of higher-order logic allows users to define equality using a single axiom,
called Leibniz equality \cite{pa-01-classical-ty-thy}:
%
$ \iforall xy. \,
   (\iforall p. \, p \, x \iimplies p \, y) \iimplies x \ieq y$.
%
Intuitively, it embodies the principle that if two terms are indistinguishable by propositions,
they must be equal.
Leibniz equality often appears in TPTP problems  \cite{gs-17-tptp}. Since modern provers have native support
for equality, it is usually beneficial to recognize and replace occurrences of Leibniz equality by the natively
supported one.

@@

\looseness=-1
Combining these two rules with dynamic clausification is more useful: When rule
\infname{IC} is replaced by rule \infname{DC}, the success rate increases
to 1660 problems, compared to 1646 problems solved by b. We also discovered
that reasoning with choice is useful: When rule \infname{Choice} is enabled, the
success rate increases to 1653. We determined that including or excluding the
conclusion $D$ of \infname{Choice}, after it is simplified, makes no difference.
Counterintuitively, disabling the \infname{BoolSimp} rule results in 1640 problems,
which is only 6 problems short of configuration b. Disabling the \infname{Abs} and
\infname{Interpret$\lambda$} rules results in solving 25 and 31 problems fewer,
respectively. The raw data show that in total, using configurations from Figure
\ref{fig:cases} to Figure \ref{fig:bool-rules}, 1682 problems can be solved.

Last, we compared our approach to alternatives. Axiomatizing Booleans brings
Zipperposition down to a grinding halt: only 1106 problems can be solved using
this mode. On the other hand, preprocessing is fairly competitive: it solved
only 8 problems fewer than the b configuration.

### 2022-07-07 conclusion/conclusion.tex
martin suda comments 2/2

While preparing for the following year's CASC competition, we found problems
that occur in practice for which the rules of \osup{} are necessary. However,
these problems occur in less than one percent of the whole TPTP library.
Furthermore, some of \osup{}'s rule are so explosive that they are disabled in
most CASC portfolio configurations.

@@

This highly pragmatic approach proved successful not only on proof assistant
benchmarks: \ehohii{} also excels on TPTP benchmarks, trailing slightly behind
Zipperposition. In the coming months we plan to replace Ehoh by \ehohii{} in Sledgehammer.
As Ehoh is already one of the most successful Sledgehammer backends,
this will further improve the efficiency of hundreds of mathematicians and computer scientists using Isabelle.
By manually expecting the TPTP benchmarks that are out of reach
for \ehohii{}, but within the reach of Zipperposition, we realized that resolving the lack of
dynamic clausification might be the weight necessary to tip the scales in
favor of \ehohii{} on TPTP benchmarks.

@@

\ourpara{Further Extensions of \ehohii{}} \ehohii{} outperforms other provers on
problems coming from proof assistants. However, its performance on TPTP problems is
slightly behind Zipperposition. We plan to further investigate the benchmarks on
which \ehohii{} fails, but that are proved by Zipperposition. One of the first
features we plan implement is dynamic clausification, which can possibly help
improve \ehohii{}'s performance on first-order problems as well. We also plan to
experiment with implementing some of the \osup{} rules necessary for
completeness such as \infname{FluidSup}. Heuristics for \ehohii{} are trained on
a subset of Sledgehammer and TPTP problems. We plan to use other benchmark sets
such as GRUNGE \cite{bgksu-19-grunge} to make \ehohii{} efficient on a wider range of higher-order
benchmarks.

\ourpara{SAT-inspired eliminations for superposition} We already implemented
some of the techniques described in Chapter~\ref{ch:satfol} in E. However, this
implementation is not properly tested and optimized. We plan to put finishing
touches to this implementation and test the effects of implemented preprocessing
techniques on E's performance. There are more techniques used in SAT solving,
notably bounded variable addition \cite{mhb-12-reencoding} and blocked clause
addition \cite{ok-99-er}, that we did not port to first-order logic. We want to
investigate whether these techniques can be lifted to first-order logic. Lastly,
we already started work on porting existing techniques to higher-order logic.

>>>>

While preparing for the following year's CASC competition, we found problems
that occur in practice for which the rules of \osup{} are necessary. However,
these problems occur in less than one percent of the whole TPTP library.
Furthermore, some of \osup{}'s rules are so explosive that they are disabled in
most CASC portfolio configurations.

@@

This highly pragmatic approach proved successful not only on proof assistant
benchmarks: \ehohii{} also excels on TPTP benchmarks, trailing slightly behind
Zipperposition. In the coming months we plan to replace Ehoh by \ehohii{} in Sledgehammer.
As Ehoh is already one of the most successful Sledgehammer backends,
this will further improve the efficiency of hundreds of mathematicians and computer scientists using Isabelle.
By manually inspecting the TPTP benchmarks that are out of reach
for \ehohii{}, but within the reach of Zipperposition, we realized that resolving the lack of
dynamic clausification might be the weight necessary to tip the scales in
favor of \ehohii{} on TPTP benchmarks.

@@

\ourpara{Further Extensions of \ehohii{}} \ehohii{} outperforms other provers on
problems coming from proof assistants. However, its performance on TPTP problems is
slightly behind Zipperposition. We plan to further investigate the benchmarks on
which \ehohii{} fails, but that are proved by Zipperposition. One of the first
features we plan to implement is dynamic clausification, which can possibly help
improve \ehohii{}'s performance on first-order problems as well. We also plan to
experiment with implementing some of the \osup{} rules necessary for
completeness such as \infname{FluidSup}. Heuristics for \ehohii{} are trained on
a subset of Sledgehammer and TPTP problems. We plan to use other benchmark sets
such as GRUNGE \cite{bgksu-19-grunge} to make \ehohii{} efficient on a wider range of higher-order
benchmarks.

\ourpara{SAT-Inspired Eliminations for Superposition} We already implemented
some of the techniques described in Chapter~\ref{ch:satfol} in E. However, this
implementation is not properly tested and optimized. We plan to put finishing
touches to this implementation and test the effects of implemented preprocessing
techniques on E's performance. There are more techniques used in SAT solving,
notably bounded variable addition \cite{mhb-12-reencoding} and blocked clause
addition \cite{ok-99-er}, that we did not port to first-order logic. We want to
investigate whether these techniques can be lifted to first-order logic. Lastly,
we already started work on porting existing techniques to higher-order logic.

### 2022-07-07 ehoh2/ehoh2.tex
martin suda comments 2/2

  \ourpara{Comparison with the First-Order E} Both Ehoh and \ehohii{} can be compiled
  in a mode that disables most of the higher-order reasoning. This mode is
  designed for users that are interested only in E's first-order capabilities and
  care a lot about performance.
  To answer the second
  evaluation question, about assessing overhead of \ehohii{},
  we chose all the \NumberOK{1138} unique problems used at CASC
  from 2019 to 2021 in the first-order theorem division and ran Ehoh and \ehohii{}
  both in this first-order (FO) mode and in higher-order (HO) mode.

>>>>

  \ourpara{Comparison with the First-Order E} Both Ehoh and \ehohii{} can be compiled
  in a mode that disables most of the higher-order reasoning. This mode is
  designed for users that are interested only in E's first-order capabilities and
  care a lot about performance.
  To answer the second
  evaluation question, about assessing overhead of \ehohii{},
  we chose all the \NumberOK{1138} unique problems used at CASC
  from 2019 to 2021 in the first-order theorem division and ran Ehoh and \ehohii{}
  both in this first-order (FO) mode and in higher-order (HO) mode using 30 s CPU and wallclock timeout.

### 2022-07-07 ho-techniques/ho-techniques.tex
martin suda comments 2/2

An earlier evaluation of the \osup{} calculus \cite{bbtv-21-full-ho-sup} showed
that the \emph{outer} variant of delayed clausification substantially increases
this calculus's performance. The outer variant clausifies top-level logical
symbols, proceeding from the outside inwards; this method corresponds to the one
described in the previous chapter. For example, a clause $C \llor
\neglit{(\cst{p} \iand \cst{q})}$ is transformed into $C \llor \neglit{\cst p}
\llor \neglit{\cst q}$. The calculus also supports \emph{inner} delayed
clausification, which uses only the core calculus rules to clausify problems.
Even though this is the laziest approach to clausification, the earlier
evaluation \cite{bbtv-21-full-ho-sup} showed that this
approach is inefficient. Thus, we focus only on the outer rules.

@@

\looseness=-1
The Boolean instantiation technique presented above requires delayed
clausification.
We assessed it in the best configuration from
%To assess it, we enabled it in the best configuration from
Figure~\ref{fig:avatar-clause}, \relax{DCS}$-$\relax{LA}. With this change ($+$\relax{BI}),
Zipperposition proves \NumberOK{1700} TPTP problems and \NumberOK{456} SH
problems.
On TPTP, even though $+$\relax{BI} solves \NumberOK{three} problems less
than \relax{DCS}$-$LA, it is very useful: \NumberOK{41}
problems can be proved with $+$\relax{BI} but not with \relax{DCS}$-$\relax{LA}. 
Conversely, \NumberOK{44} problems are solved with \relax{DCS}$-$\relax{LA},
but not with $+$\relax{BI}, which suggests that Boolean instantiation can be 
explosive.
%
One of the problems Boolean instantiation helps solve is \texttt{NUM636\^{}2} (a
reencoding of \texttt{NUM636\^{}3}, which is discussed in Example \ref{hot:exa:num636}).
It conjectures that $\iforall x.\, \cst{s} \, x \ineq x$, where $x$ ranges over
Peano-style numbers specified by $\cst{z}$ and $\cst{s}$. The given axioms are
the induction principle $\iforall p.\, p \, \cst{z} \,\iand\, \iforall x. \, (p \,
x \iimplies p \, (\cst{s} \, x)) \iimplies \iforall x. \, p\, x$, injectivity
$\iforall x y. \, \cst{s}\,x \ieq \cst{s}\,y \iimplies x \ieq y$, and
distinctness $\iforall x. \, \cst{s}\,x \ineq \cst{z}$. The conjecture is easily
proved if Boolean instantiation is enabled: Even though the conjecture literal
cannot be abstracted, instantiating $p$ with the term $\lambda x.\, \cst{s}
\, x \ineq x$ used in the encoding of the (nonclausified) conjecture leads to a proof in just
\NumberOK{22} given clause loop iterations. Zipperposition also finds a
proof using the \relax{DCI}$-$\relax{LA} configuration, but this requires
\NumberOK{294} iterations.

@@

\begin{exa} The green subterms of the term $F \, \cst{a} \, \iand \, \cst{p} \,
(\iforall \, (\lambda x. \, \cst{q} \, x)) \, \cst{b}$ are the term itself, $F \, \cst{a}$, $\cst{p} \, (\iforall \, (\lambda x. \, \cst{q} \, x))
\, \cst{b}$, $\iforall \, (\lambda x. \, \cst{q} \, x)$, and $\cst{b}$.
\end{exa} 
%
Green positions are lifted to clauses as follows: If $p$ is the green position
of a subterm in $s$, and $s$ occurs in a literal $l \in \{s \eq t{,}\; s \not\eq
t\}$ of a clause $C$, the green position of the same subterm in $C$ is denoted by $l.s.p$.
\osup{} mandates additional restrictions on the Boolean selection function:
$\itrue$, $\ifalse$, and variable-headed terms cannot be selected; for literals
$s \eq t$, neither $s$ nor $t$ cannot be selected; if $s$ (or $t$) contains a
variable $X$ as a green subterm and $X\> \tuplen{u}$, with $n \ge 1$, is a
maximal term of $C$, then $s$ (or $t$) cannot be selected.

@@

Ganzinger and Stuber considered Boolean subterm selection for their extension of
first-order superposition with interpreted Boolean type \cite{gs-05-boolsup}. Unlike our calculus, their calculus requires
only allows the selection of subterms that occur in negative green positions, defined below.

@@

\begin{exa} 
  \begin{sloppypar}
  In this example we simplify the notation by writing positive predicate literals
  $\poslit{s}$ as $s$ and negative predicate literals $\neglit{t}$ as $\neg t$.
  Consider the unsatisfiable two-clause problem $\{ X \, (\cst{f} \,
  \cst{a}) \not\eq \cst{f} \, (X \, \cst{a}) \lor \cst{p} \, (X \, \cst{a}),\allowbreak
  \neg \cst{p} \, (\cst{f}^{100} \, \cst{a})  \}$ and a selection function which
  selects negative literals. 
  Let $P \mid A \mid Q$ denote
  the state of the given clause loop (i.e., the contents of the passive and active set
  and of the stream queue), and let $\infstream{ a_1, a_2, \ldots }$
  denote an infinite stream of elements.
  \end{sloppypar}

@@

% TODO: If I add a new benchmark set, make sure this sentence is fixed
\looseness=-1
To evaluate our unification algorithm (using a somewhat different
evaluation setup and an older version of Zipperposition), we also compared a complete, nonterminating variant of the
unification procedure and a pragmatic, terminating variant. The
pragmatic variant was used directly---all the inference conclusions were put
immediately in $P$, bypassing $Q$. The complete variant, which relies on
possibly infinite streams and is much more prolific, proved only 15  problems
less than the most competitive pragmatic variant. Furthermore, it proved 19
problems not proved by the pragmatic variant.
%
This shows that our given clause procedure, with its heuristics, allows the
prover to defer exploring less promising branches of the unification and uses
the full power of a complete higher-order unifier search to solve unification
problems that cannot be proved by a restricted procedure.

>>>>

An earlier evaluation of the \osup{} calculus \cite{bbtv-21-full-ho-sup} showed
that the \emph{outer} variant of delayed clausification substantially increases
this calculus's performance. The outer variant clausifies top-level logical
symbols, proceeding from the outside inwards; this method corresponds to the one
described in the previous chapter. For example, a clause $C \llor
\neglit{(\cst{p} \iand \cst{q})}$ is transformed into $C \llor \neglit{\cst p}
\llor \neglit{\cst q}$. The calculus also supports \emph{inner} delayed
clausification, which does not enforce clausifying logical symbols in top-down direction, 
and uses only the generating calculus rules to clausify problems.
Even though this is the laziest approach to clausification, the earlier
evaluation \cite{bbtv-21-full-ho-sup} showed that this
approach is inefficient. Thus, we focus only on the outer rules.

@@

\looseness=-1
The Boolean instantiation technique presented above requires delayed
clausification.
We assessed it in the best configuration from
%To assess it, we enabled it in the best configuration from
Figure~\ref{fig:avatar-clause}, \relax{DCS}$-$\relax{LA}. With this change ($+$\relax{BI}),
Zipperposition proves \NumberOK{1700} TPTP problems and \NumberOK{456} SH
problems.
On TPTP, even though $+$\relax{BI} solves \NumberOK{three} problems fewer
than \relax{DCS}$-$LA, it is very useful: \NumberOK{41}
problems can be proved with $+$\relax{BI} but not with \relax{DCS}$-$\relax{LA}. 
Conversely, \NumberOK{44} problems are solved with \relax{DCS}$-$\relax{LA},
but not with $+$\relax{BI}, which suggests that Boolean instantiation can be 
explosive.
%
One of the problems Boolean instantiation helps solve is \texttt{NUM636\^{}2} (a
reencoding of \texttt{NUM636\^{}3}, which is discussed in Example \ref{hot:exa:num636}).
It conjectures that $\iforall x.\, \cst{s} \, x \ineq x$, where $x$ ranges over
Peano-style numbers specified by $\cst{z}$ and $\cst{s}$. The given axioms are
the induction principle $\iforall p.\, p \, \cst{z} \,\iand\, \iforall x. \, (p \,
x \iimplies p \, (\cst{s} \, x)) \iimplies \iforall x. \, p\, x$, injectivity
$\iforall x y. \, \cst{s}\,x \ieq \cst{s}\,y \iimplies x \ieq y$, and
distinctness $\iforall x. \, \cst{s}\,x \ineq \cst{z}$. The conjecture is easily
proved if Boolean instantiation is enabled: Even though the conjecture literal
cannot be abstracted, instantiating $p$ with the term $\lambda x.\, \cst{s}
\, x \ineq x$ used in the encoding of the (nonclausified) conjecture leads to a proof in just
\NumberOK{22} given clause loop iterations. Zipperposition also finds a
proof using the \relax{DCI}$-$\relax{LA} configuration, but this requires
\NumberOK{294} iterations.

@@

\begin{exa} The green subterms of the term $F \, \cst{a} \, \iand \, \cst{p} \,
(\iforall \, (\lambda x. \, \cst{q} \, x)) \, \cst{b}$ are the term itself, $F \, \cst{a}$, $\cst{p} \, (\iforall \, (\lambda x. \, \cst{q} \, x))
\, \cst{b}$, $\iforall \, (\lambda x. \, \cst{q} \, x)$, and $\cst{b}$.
\end{exa} 
%
Green positions are lifted to clauses as follows: If $p$ is the green position
of a subterm in $s$, and $s$ occurs in a literal $l \in \{s \eq t{,}\; s \not\eq
t\}$ of a clause $C$, the green position of the same subterm in $C$ is denoted by $l.s.p$.
\osup{} mandates additional restrictions on the Boolean selection function:
$\itrue$, $\ifalse$, and variable-headed terms cannot be selected; for literals
$s \eq t$, neither $s$ nor $t$ can be selected; if $s$ (or $t$) contains a
variable $X$ as a green subterm and $X\> \tuplen{u}$, with $n \ge 1$, is a
maximal term of $C$, then $s$ (or $t$) cannot be selected.

@@

Ganzinger and Stuber considered Boolean subterm selection for their extension of
first-order superposition with interpreted Boolean type \cite{gs-05-boolsup}. Unlike our calculus, their calculus
only allows the selection of subterms that occur in negative green positions, defined below.

@@

\begin{exa} 
  \begin{sloppypar}
  In this example we simplify the notation by writing positive predicate literals
  $\poslit{s}$ as $s$ and negative predicate literals $t \noteq \itrue$ as $\neg t$.
  Consider the unsatisfiable two-clause problem $\{ X \, (\cst{f} \,
  \cst{a}) \not\eq \cst{f} \, (X \, \cst{a}) \lor \cst{p} \, (X \, \cst{a}),\allowbreak
  \neg \cst{p} \, (\cst{f}^{100} \, \cst{a})  \}$ and a selection function which
  selects negative literals. 
  Let $P \mid A \mid Q$ denote
  the state of the given clause loop (i.e., the contents of the passive and active set
  and of the stream queue), and let $\infstream{ a_1, a_2, \ldots }$
  denote an infinite stream of elements.
  \end{sloppypar}

@@

% TODO: If I add a new benchmark set, make sure this sentence is fixed
\looseness=-1
To evaluate our unification algorithm (using a somewhat different
evaluation setup and an older version of Zipperposition), we also compared a complete, nonterminating variant of the
unification procedure and a pragmatic, terminating variant. The
pragmatic variant was used directly---all the inference conclusions were put
immediately in $P$, bypassing $Q$. The complete variant, which relies on
possibly infinite streams and is much more prolific, proved only 15  problems
fewer than the most competitive pragmatic variant. Furthermore, it proved 19
problems not proved by the pragmatic variant.
%
This shows that our given clause procedure, with its heuristics, allows the
prover to defer exploring less promising branches of the unification and uses
the full power of a complete higher-order unifier search to solve unification
problems that cannot be proved by a restricted procedure.

### 2022-07-07 satfol/satfol.tex
martin suda comments 2/2

The last technique we study is \emph{blocked clause elimination} (Sect.\
\ref{sec:satfol:satisfiability-by-clause-elimination}). It is used in both SAT
\cite{jbh-10-BCE} and QBF solvers \cite{bls-11-bloqqer}. Its generalization to first-order
logic has produced good results when used as a preprocessor, especially on
satisfiable problems \cite{ksstb-2017-blockedfol}. We explore more ways to use
blocked clause elimination on satisfiable problems, including using it to
establish equisatisfiability with an empty clause set or as an inprocessing
rule. Unfortunately, we find that its use as inprocessing can compromise the
refutational completeness of superposition.

@@

\end{rep}
\looseness=-1
For first-order logic without equality, a clause $L \cor C$ is blocked if all its
$L$-resolvents are tautologies \cite{ksstb-2017-blockedfol}. The $L$-resolvent
between $L \cor C$ and $\neglit{L_1} \cor \cdots \cor \neglit{L_n} \cor
D$ (which are renamed apart) is $\sigmacl{C \cor D}$, where $\sigma$ is the MGU
of the literals $L, L_1, \ldots, L_n$.
Given a Herbrand model $\ourmodel$ of a problem, the following procedure
removes all clauses while preserving satisfiability:

@@

  \Procedure{AddImplication}{$L_\mathrm{a}{,}\> L_\mathrm{c}{,}\> C$}
    \If{$\impmap[\sigmaterm{L_\mathrm{a}}] \not = \emptyset$ for some renaming $\sigma$} \label{statement:alpha}
      \State $(L_\mathrm{a}, L_\mathrm{c}) \gets (\sigmaterm(L_\mathrm{a}), \sigmaterm{L_\mathrm{c}})$
    \EndIf

>>>>

The last technique we study is \emph{blocked clause elimination} (Sect.\
\ref{sec:satfol:satisfiability-by-clause-elimination}). It is used in both SAT
\cite{jbh-10-BCE} and QBF solvers \cite{bls-11-bloqqer}. Its generalization to first-order
logic has produced good results when used as a preprocessor, especially on
satisfiable problems \cite{ksstb-2017-blockedfol}. We explore more ways to use
blocked clause elimination on satisfiable problems, including using it to
establish equisatisfiability with the empty clause set or as an inprocessing
rule. Unfortunately, we find that its use as inprocessing can compromise the
refutational completeness of superposition.

@@

\end{rep}
\looseness=-1
For first-order logic without equality, a clause $L \cor C$ is blocked if all its
$L$-resolvents are tautologies \cite{ksstb-2017-blockedfol}. The $L$-resolvent
between $L \cor C$ and $\neglit{L_1} \cor \cdots \cor \neglit{L_n} \cor
D$ (renamed apart) is $\sigmacl{C \cor D}$, where $\sigma$ is the MGU
of the literals $L, L_1, \ldots, L_n$.
Given a Herbrand model $\ourmodel$ of a clause set $N$, the following procedure
removes all clauses while preserving satisfiability:

@@

  \Procedure{AddImplication}{$L_\mathrm{a}{,}\> L_\mathrm{c}{,}\> C$}
    \If{$\impmap[\sigmaterm{L_\mathrm{a}}] \not = \emptyset$ for some renaming $\sigma$} \label{statement:alpha}
      \State $(L_\mathrm{a}, L_\mathrm{c}) \gets (\sigmaterm{L_\mathrm{a}}, \sigmaterm{L_\mathrm{c}})$
    \EndIf

### 2022-07-07 unif/unif.tex
spellcheck
martin suda comments 2/2

of the variables with respect to the free variables ocurring in $E$. The transitions, defined as follows, are only

@@

      $\overline{z}_m$ are bound variables fresh with respect to $s$ and $t$, $s' = \{x_1 \mapsto z_1, \ldots, x_m \mapsto z_m \}$,
      and $t' = \{y_1 \mapsto z_1, \ldots, y_n \mapsto z_n \}$

@@

\section{Evaluation}
\label{sec:unif:evaluation}We evaluated the implementation of our unification procedure in Zipperposition,
%TODO: @PETAR: "the complete and the pragmatic variant"? 
%Done.
% Also, the complete variant is sometimes called "full". Maybe we should stick to one of the two?
% @ALEX: Investingating.
assessing the complete variant and the pragmatic variant, the latter with several
different combinations of limits for number of bindings. As part of the
implementation of the complete mode for Boolean-free higher-order logic in
Zipperposition \cite{bbtvw-21-sup-lam}, Bentkamp implemented a straightforward
version of the JP procedure. This version is faithful to the original description,
with a check as to whether a (sub)problem can be solved using a first-order oracle
as the only optimization. Our evaluations were performed on StarExec Miami
\cite{sst-14-starexec} servers with Intel Xeon E5-{2620 v4} CPUs clocked at {2.10}\,GHz with 60\,s CPU limit.

>>>>

of the variables with respect to the free variables occurring in $E$. The transitions, defined as follows, are only

@@

      $\overline{z}_m$ are bound variables fresh with respect to $s$ and $t$, $s' = \{x_1 \mapsto z_1, \ldots, x_m \mapsto z_m \}\,s$,
      and $t' = \{y_1 \mapsto z_1, \ldots, y_n \mapsto z_n \}\,t$

@@

\section{Evaluation}
\label{sec:unif:evaluation}We evaluated the implementation of our unification procedure in Zipperposition,
%TODO: @PETAR: "the complete and the pragmatic variant"? 
%Done.
% Also, the complete variant is sometimes called "full". Maybe we should stick to one of the two?
% @ALEX: Investigating.
assessing the complete variant and the pragmatic variant, the latter with several
different combinations of limits for number of bindings. As part of the
implementation of the complete mode for Boolean-free higher-order logic in
Zipperposition \cite{bbtvw-21-sup-lam}, Bentkamp implemented a straightforward
version of the JP procedure. This version is faithful to the original description,
with a check as to whether a (sub)problem can be solved using a first-order oracle
as the only optimization. Our evaluations were performed on StarExec Miami
\cite{sst-14-starexec} servers with Intel Xeon E5-{2620 v4} CPUs clocked at {2.10}\,GHz with 60\,s CPU limit.
